---
title: "Creating the Interactive User Input for Automating the Dashboard"
output: rmarkdown::html_vignette
author: "Cory Roth"
vignette: >
  %\VignetteIndexEntry{Creating the Interactive User Input for Automating the Dashboard}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, include=FALSE}
#library(DSPGFarmer)
```

This is a php script used with OutlineUI.html for DSPG Farmer-Asset-Mapping

Allows Users to select a bounding box where their farm is, along wimth inputs for weather data they want

will take inputs and run R scripts based on those inputs to produce CSV's of data to be used in Tableau

Current Time it takes ~2minutes(30-45sec for weather and 1:15 for soil)

# PHP code

In our HTML file, the user will zoom into the region they want data for, and give use other inputs

1. Year of weather data to go back to
2. Range of Months to look at for weather

As they submit their inputs, HTML posts them to PHP, where they are received and assigned to variables
```{eval=FALSE}
header('Access-Control-Allow-Origin *');
 
#coords should be an array of 4 numbers, xmin,xmax,ymin,ymax of a bounding box

$coords = $_POST['coords'];


#startmonth is for range of months to look at for weather data

$startmonth = $_POST['smonth'];


#endmonth is for range of months to look at for weather data

$endmonth = $_POST['emonth'];


#year is for weather data

$year = $_POST['year'];

```


We take the coords bounding box and assign each extreme a value
```{eval=FALSE}
    $xmin = $coords[0];
    $xmax = $coords[1];
    $ymin = $coords[3];
    $ymax = $coords[2];
```


### Centroid
For the weather data, we want to be using the center of the Bounding Box, so we run an R script using command exec() to run a command line call to run the R script to calculate the centroid, given the bounding box as inputs and recieving the x,y coords pair as outputs

```{eval=FALSE}
$centroid = exec("\"C:\\Program Files\\R\\R-4.2.0\\bin\\Rscript.exe\" C:\\xampp\\htdocs\\centroid_calc.R $xmin $xmax $ymin $ymax");
```

The return value is a string in the format "x,y" of coordinates representing the centroid, we want to convert this string into an array

```{eval=FALSE}
	$centpoints = preg_split("/\,/", $centroid);
```

## RScripts

  Now that we have all the inputs needed to obtain data
  
  We run our R code using Rscript
  
  The first script ran is the script that obtains the raw weather data from IEMRE
  
  5 inputs: year, startmonth, endmonth, and centerpoints
   
  This does not need to return anything to PHP because it writes its data straight to a csv locally
  
```{eval=FALSE}
	exec("\"C:\\Program Files\\R\\R-4.2.0\\bin\\Rscript.exe\" \"C:\\Users\\cornd\\OneDrive\\Documents\\GitHub\\Farmer-Asset-Mapping\\Data Exploration\\TestPHP\\WeatherScript.R\" $year $startmonth $endmonth $centpoints[0] $centpoints[1]");

```


 After that is ran, we can run the script that aggregates the weather data and the script that calculated GDD for each crop
 
 We have to run these after the weather scrapping script because they are dependent on the raw weather data 
 
 They do not need inputs because they will be using the csv as their input source
 
 They both save their data as local csv's
 
```{eval=FALSE}
	exec("\"C:\\Program Files\\R\\R-4.2.0\\bin\\Rscript.exe\" \"C:\\Users\\cornd\\OneDrive\\Documents\\GitHub\\Farmer-Asset-Mapping\\Data Exploration\\TestPHP\\GDD.R\"");
	
	exec("\"C:\\Program Files\\R\\R-4.2.0\\bin\\Rscript.exe\" \"C:\\Users\\cornd\\OneDrive\\Documents\\GitHub\\Farmer-Asset-Mapping\\Data Exploration\\TestPHP\\TempDataSummary.R\"");

```
 
Finally We can run our Soil gathering script

This script obtains the soil data and merges it with our specialty crop data

It also obtain the county FIPS code it is in and the State ID to use in updating the Risk Managment Url querys

6 inputs: bounding box,  centerpoints

Saves data as a CSV
```{eval=FALSE}
	exec("\"C:\\Program Files\\R\\R-4.2.0\\bin\\Rscript.exe\" \"C:\\Users\\cornd\\OneDrive\\Documents\\GitHub\\Farmer-Asset-Mapping\\Data Exploration\\TestPHP\\Cory_CropInfoV3.R\" $xmin $xmax $ymin $ymax $centpoints[0] $centpoints[1]");

```


After all data has been collected and stored

PHP will output to HTML that the data has been collected

```{eval=FALSE}
	echo "Data Collected";
```

