---
title: "Retreiving and Connecting Soil Data to the Dashboard"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Retreiving and Connecting Soil Data to the Dashboard}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The libraries Used To Gather, Clean and Merge Soil Data

```{r setup, eval =FALSE}
#library(DSPGFarmer)
library(FedData)
library(readxl)
library(dplyr)
library(tidyr)
library(sqldf)
library(raster)
```



Gathering Inputs from PHP

```{r, eval=FALSE}
setwd("C:\\Users\\cornd\\OneDrive\\Documents\\GitHub\\Farmer-Asset-Mapping\\Data Exploration\\TestPHP")
args <- commandArgs(TRUE)

##Bounding Box Input
s <- raster(ncol=2, nrow=2)

##INPUTS from php
ymin(s) <- as.numeric(args[1])
ymax(s) <- as.numeric(args[2])
xmin(s) <- as.numeric(args[3])
xmax(s) <- as.numeric(args[4])

lat = as.numeric(args[5])
lon= as.numeric(args[6])

```

# Gather Data from Web Soil Survey (SSURGO Data)
### Input:
  can Either be a Rasterlayer or a Spatial object
  
  For this, we are using a Rasterlayer that uses the coordinates selected from the user as the bounding box
  
### Output:
  It downloads 2 Folders to local File, a __RAW__ and an __EXTRACTIONS__ folder
  
  Raw is raw zipped data, EXTRACTIONS holds info being used, has both tabular (.csv) and spatial object (.shp)
  
  In Extraction, will have folder named as label parameter
  

  
  
### force.redo:
  Parameter in function that if
  
  set to false, means that if there is already data in label folder, then it won't download data from web soil survey
  
  set to true, any time it is called it will download data from web soil survey and overwrite what is stored in label folder
  
```{r, eval=FALSE}
Area<- get_ssurgo(template = s,label = "CropSelection_V3", force.redo =TRUE)
```
  
# Cleaning and Merging SSURGO Data

The SSURGO data has both spatical and tabular data, but for cleaning we only want to look at the tabluar

Will Merge it to the spatial file in tableau

[More Information About SSURGO data structure](https://www.nrcs.usda.gov/wps/portal/nrcs/detail/soils/survey/?cid=nrcs142p2_053627)

```{r, eval=FALSE}
Data <- Area$tabular

##Tables of data we are using
componet <- Data$component
muaggart <-Data$muaggatt
mapunit <-Data$mapunit
comonth <- Data$comonth %>% filter(month=="May") ##only want month of may to follow CSR2
coerosionacc <- Data$coerosionacc %>% filter(rvindicator=="Yes") ##only care about dominant areas
chorizon <- Data$chorizon
chtexturegrp <-Data$chtexturegrp %>%filter(rvindicator == "Yes") ##only care about dominant areas
```


Merging The map units data and components together
```{r, eval=FALSE}
Overall<- merge(muaggart,componet,by = "mukey", all = TRUE)
Overall<- merge(Overall,mapunit,by = "mukey", all = TRUE)
Overall<- merge(Overall,comonth,by = "cokey", all = TRUE)
Overall<- merge(Overall,coerosionacc,by = "cokey", all = TRUE)

chorizon2 <- merge(chorizon,chtexturegrp, by = "chkey", all.x=TRUE)
chorizon2 <-chorizon2%>%
  mutate(depth = hzdept.r /2.54) ##convert cm to inches
```


Some Data of SSURGO comes in at varying depth levels, like pH or Available Water Storage

Because Crops have a specific max rooting depth, we only want to look at the values of data that are in the rooting depths range
(0 - Max Rooting Depth)

CropData is our Database of specialty crops that includes the rooting depth

sqldf (which uses SQL) is used because we are merging based on an __inequality__ and it makes it easier to do than in R

```{r, eval=FALSE}
CropData <- read_excel("Input\\Crop-Info_Farmer Asset Mapping.xlsx")


##Merge Soil Data With Crop Data, gets rid of all values that are past the max Rooting Depth (Later going to average depths)
MergedData <- sqldf("select * from chorizon2 left join CropData
             on (chorizon2.depth <= CropData.Depth_l)")

##Filters for areas without crops
MergedDataNoNA <- MergedData %>%
  filter(!is.na(`Types of Crops`))

```

Now that we are looking at the right depths, we find what value to use

For quantitative values like pH and AWS:
  
    We use the average based on all depths in range
    
For qualitative values like soil texture:
  
    We follow how the query tool on the Web soil survey does and use the surface layer data



```{r, eval=FALSE}
chorzonSim <-MergedDataNoNA%>%
  mutate(depthLevel = "0-MaxRootDepth")

##for values that very on depth, chose to take average of or value on surface layer
chorzonSim2<- chorzonSim%>%
  filter(depthLevel=="0-MaxRootDepth")%>%
  group_by(cokey,`Types of Crops`)%>%
  mutate(soilText= ifelse(depth==0,texdesc,""))%>%
  summarise(ksat = mean(ksat.r,na.rm=TRUE),awc = mean(awc.r, na.rm=TRUE),caco3 = mean(caco3.r,na.rm=TRUE),gypsum= mean(gypsum.r,na.rm=TRUE),sar =mean(sar.r,na.rm=TRUE),ec = mean(ec.r,na.rm=TRUE),cec7= mean(cec7.r,na.rm=TRUE), ph = mean(ph1to1h2o.r,na.rm=TRUE),ph_l = mean(ph1to1h2o.l,na.rm=TRUE),ph_h= mean(ph1to1h2o.h,na.rm=TRUE),om= mean(om.r,na.rm=TRUE),
            ptotal = mean(ptotal.r,na.rm=TRUE),soilTextsum = soilText,Kfact = mean(kffact,na.rm=TRUE))%>%
  filter(soilTextsum!="")
```

Now that data is averaged, merge data with crop data

```{r,eval=FALSE}
##Now that data is averaged, remerge data with crop data
chData <- merge(chorzonSim2,CropData,by = "Types of Crops")

Overall2<- merge(Overall,chData,by = "cokey", all = TRUE)
Overall2 <- Overall2 %>%
  filter(majcompflag=="Yes") ##Use the componets that are the majorty of the mapunit
```

Now that everything is all together, we select only the values that we want to use

```{r, eval= FALSE}
Simple <- as_tibble(Overall2) %>%
  dplyr::select(musym.x,mukey,cokey,muname = muname.x,taxorder,compname,slope.r,slopegradwta, slope.l,slope.h,localphase,erokind,erocl,tfact,Kfact,wei,
         niccdcd,hydgrp,soilslippot,drainagecl,drclassdcd,niccdcd,awc,aws025wta,aws0150wta, flodfreqcl,floddurcl,pondfreqcl,
         ponddurcl,flodfreqdcd,flodfreqmax,pondfreqprs,iacornsr,ph,ph_l,ph_h,cec7,gypsum,ksat,ec,sar,caco3,om,ptotal,
         soilTextdes =soilTextsum,`Types of Crops`,`Soil Types`,`Rooting Depth`,`pH-Level`,`Temperature Tolerances`,ph_L,ph_H,Boron,Copper,Zinc,Molybdenum,Iron,Manganese,`Soil Boron`,`Soil Magnesium`,`Storage Life`,`Crop Maturity Late Variety`,`Crop Maturity Early Variety (days)`,`Storage Humidity (%)`,`Storage Temp (  ̊F)`,`Base Temp in F`,Nitrogen)
```

In order for our crop rotation dashboard to be linked with our crop data, we will merge it together here

```{r, eval=FALSE}

Simple <- merge(Simple,CropRotation, by.x = "Types of Crops", by.y = "Crop", all.x=TRUE )

```


## Adding Flags

This section of code is going to be how we determine flags

Based on the corn sutability rating and the information we gleamed from research, we have a list of 10 variables that deal with soil and are important to crop growth 

We will determine how many issues there are for all map units in the selected area for __each__ crop

```{r , eval=FALSE}
Simple<- Simple%>%
  mutate(Flags =
           ifelse(ph <= ph_H & ph>=ph_L,0,1),FlagDesc = ifelse(ph <= ph_H & ph>=ph_L,"",",PH does not fit into range"))%>%
  mutate(Flags =
           ifelse(soilTextdes != `Soil Types`,Flags+1,Flags),FlagDesc = ifelse(soilTextdes!=`Soil Types`,paste(FlagDesc,"soil texture does not match",sep=', '),FlagDesc))%>%
  mutate(Flags =
           ifelse(tfact <4,Flags+1,Flags),FlagDesc =ifelse(tfact <4,paste(FlagDesc,"this soil's erosion tolerance may be an issue",sep=', '),FlagDesc) )%>%
  mutate(Flags  =
           ifelse(Kfact> .38,Flags+1,Flags),FlagDesc =ifelse(Kfact> .38,paste(FlagDesc,"this soil is more susceptible to erosion",sep=', '),FlagDesc))%>%
  mutate(Flags  =
           ifelse(om<2 | om>8,Flags+1,Flags),FlagDesc =ifelse(om< 2|om>8,paste(FlagDesc,"this soil's organic matter percent could be an issue",sep=', '),FlagDesc))%>%
  mutate(Flags  =
           ifelse(erocl == "Class 2",Flags+1,Flags),FlagDesc =ifelse(erocl == "Class 2",paste(FlagDesc,"this soil's topsoil may have been depleated",sep=', '),FlagDesc))%>%
  mutate(Flags  =
           ifelse(aws0150wta <=9.00,Flags+1,Flags),FlagDesc =ifelse(aws0150wta<=9.00,paste(FlagDesc,"this soil's water holding may be limited",sep=', '),FlagDesc))%>%
  mutate(Flags  =
           ifelse(drclassdcd =="Very poorly drained" | drclassdcd == "Poorly drained",Flags+1,Flags),FlagDesc =ifelse(drclassdcd =="Very poorly drained" | drclassdcd == "Poorly drained",paste(FlagDesc,"this soil's drainage may be limited",sep=', '),FlagDesc))%>%
  mutate(FlagDesc =  ifelse(substr(FlagDesc,1,1)==',',substr(FlagDesc,3,nchar(FlagDesc)),FlagDesc))
```


To save having to output the fips code and state abbreviation back into php and run this script, we just run it from this R script 

This Code takes in the lat and lon of the user, along with the County FIPS code and State ID to update the links we provide so that when users select the links, it shows to the location they selected

```{r, eval=FALSE}
##State Fips and State Abv (IA)
##Used In Updates Risk URL PARAMETERS
Statefips<- as.numeric(sum(19000,as.numeric(substr(Data$legend$areasymbol[1],3,5))))
StateAbv <-substr(Data$legend$areasymbol[1],0,2)
source("Cory_UpdateRiskURLParams.R")

```


Finally, we save our data to a csv
```{r, eval=FALSE}
write.csv(Simple,"Output\\CropSelection2.csv", row.names=FALSE)
```
